Equilibria and Convergence of Fictitious Play on Network Aggregative Games

### 2021-05-20 Documents/FP in NAGs/neurips_2021.tex

\title{Fictitious Play in Network Aggregative Games}

@@

\begin{abstract}
  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
  both the left- and right-hand margins. Use 10~point type, with a vertical
  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
  bold, and in point size 12. Two line spaces precede the abstract. The abstract
  must be limited to one paragraph.
\end{abstract}

@@

Multi-Agent Learning requires a number of agents to adapt in response
to the behaviour of the other agents as well as the environment
\cite{Schwartz}. This feature leads to a fundamentally non-stationary
problem, which presents a challenge to designing effective learning
policies. Even for a small number of agents in the game,
learning has been shown to lead to non-stationary, even chaotic
behaviour \cite{SatoChaos}, and this problem becomes even more
pronounced as the number of agents increase \cite{Sanders}. In this
light, it would seem that, when there is a large population of agents,
it is almost impossible to understand the long term behaviour of multi-agent learning \cite{PiliourasChaoticMaps}. This, of course, poses a
problem for AI safety, for which it is \fb{desirable} that learning systems
are designed so that their ultimate behaviour is guaranteed to satisfy
some predetermined goals.

\fb{if space allows, you might consider to contrast the situation with
  single-agent learning.}

A promising approach to solve multi-agent learning
%To deal with this problem, one must attempt
consists in reducing the many player game to something which is more
tractable. A number of reduction methods have been proposed in recent
years, most notably \emph{mean field} games \cite{} and \emph{aggregative}
games \cite{}. The former make the assumption of an infinite number of
agents, so that the population can be represented through a
distribution over players' states \cite{CainesPaper} . Every agent
then updates their action profiles depending on this distribution. In
the latter each agent plays \fb{?} considers a real valued function which is
a convex combination of the states of the other agents. In both
formulations, a many-player game is reduced to the set of two-player
games which each agent is involved in \fb{?}.

Both approaches have been the object of rigorous study and it has been
shown that agents reach an equilibrium when learning on such games
(cf.~\cite{MFGLearningPapers} for mean field games and
\cite{Aggregative Papers} for aggregative games). However, both also
present a fundamental limitation.  Namely, they both require that
agents have access to the action profiles of the entire population.
This could be through communication with all other agents, or through
the intervention of a central coordinator who is able to \fb{?} the
entire population. Whilst recent work aims to relax this assumption
through the introduction of noise \cite{MFG-FP} or partial
observability \cite{AAMASPaper}, the requirement that each agent
updates their actions based on the entire population is rather strong
\fb{and not always supported by empirical evidence}.

In this study, we investigate a variant of aggregative games:
%known as the
\emph{network aggregative} (NA) games. This format assumes that
there is an underlying communication network through which agents
interact. Then each agent updates their actions according only to
those agents with whom they communicate. This assumption significantly relaxes
the communication load on each agent and lifts the need for a central
coordinator. Recent work on NA games has shown that it is possible
for agents to reach an equilibrium strategy in an entirely distributed
manner \cite{Grammatico, LeaderFollower, MyopicAgents}. We contribute
in this direction by analysing the long term behaviour of multi-agent learning on
 NA games. In particular, we analyse the fictitious play learning
algorithm \cite{Brown, Harris}, in which agents are assumed to be
myopic, in that they react solely to the previous behaviour of the
others.

\fb{maybe above we can reduce a bit the discussion of NA and motivate
  a bit more FP.}

@@

  The main contribution of this work is to introduce learning in
  network aggregative game through the action of fictitious play (FP).

  We first show that \fb{in aggregative games played with FP} a Nash Equilibrium exists and that FP admits
  solutions in this setting. Specifically, we consider zero-sum games
  and show that FP converges to a fixed point, which for networks
  without self-loops corresponds to a Nash equilibrium. In
  addition, we find that, for games which are not zero-sum, agents
  following FP are able to achieve no regret.

  Further, we explore FP through numerical simulations to check
%onsider the question of
whether it converges in non-zero sum games. We answer this in the
negative by finding a family of games in which action profiles cycle
around the Nash equilibrium. Finally, our experiments document how
noise affects the convergence of FP, suggesting
%Our experiments suggest
that,
under the presence of noise, the algorithm still reaches a fixed
point, but perhaps not the Nash equilibrium. This presents an
interesting avenue for future research.

@@

NA games are a recent \cite{Parise2015} extension of
%introduction  which presents
%a variant of
aggregative games by adding an underlying structure to
the agent population. Since its introduction, distributed algorithms have
been built with the aim of finding NE in NA games. In particular,
\cite{Parise2015,Parise2020} consider the case in which payoffs are
given by Lipschitz functions with unique minimisers and apply standard
topological fixed-point towards designing algorithms that
converge to NE. We note that the \emph{Mann Iteration}, one of the
algorithms applied in \cite{Parise2020} shares a remarkably similar
form with the discrete variant of FP. To the best of our knowledge,
this link has not been explored in the literature and is an avenue for
research which we are currently pursuing. Another common algorithm for
distributed NE seeking is the projected gradient (resp. subgradient)
dynamics which is explored in \cite{Zhang2020}
(resp. \cite{Shokri2020, Shokri2021}). In all of these works, the cost
function is assumed to be convex, and therefore have a unique
minimiser. In fact this is a common assumption in works concerning NA
Games \cite{Zhu2021, Lei2020}, which we believe is due to its ubiquity
in control settings. We have not yet come across works which consider
NA games from the point of view of payoff matrices, which are more
common in multi agent learning settings.  Furthermore, to the best of
our knowledge, this is the first work which introduces the application
of a strict learning algorithm in the NA setting.

@@

  \subsection{Existence of the Nash Equilibrium}
  \label{sec::ExistenceofNE}

@@

  To show existence of an NE we will need the following definitions which can be found in most texts on topology \cite{Munkres}.

  \begin{definition}[Convexity]
  	A set $\mathcal{X}$ is \emph{convex} if, for every $x, y \in \mathcal{X}$ and every $\lambda \in [0, 1]$ we have that $\lambda x + (1 - \lambda)y \in \mathcal{X}$.
  \end{definition}

	\begin{definition}[Convex Combinations]
		The point $\sum_{i = 1}^{n} \lambda_i x_i$ is a \emph{convex combination} of the points $x_1, \ldots, x_n$ if $\lambda_i \geq 0$ for all $i$ and $\sum_i \lambda_i = 1$.
	\end{definition}

	\begin{definition}[Unit Simplex]
		The $n$-dimensional \emph{unit simplex} is defined as the convex combination of the $n$ standard basis vectors $(e_i)_{i = 1}^n$ in $\mathbb{R}^n$ in which $e_i$ has $1$ in the $i$'th component and $0$ in all others. In particular this is given by

		\begin{equation}
			\Delta = \{x \in \mathbb{R}^n : x = \sum_{i = 1}^n \lambda_i e_i \text{ for some } \lambda_1, \ldots, \lambda_n \geq 0, \sum_i \lambda_i = 1\}
		\end{equation}
	\end{definition}

	It follows from the definitions, then, that the unit simplex is a convex set. In addition we have that the simplex is \emph{closed} and \emph{bounded} in $\mathbb{R}^n$. By the Heine-Borel Theorem \cite{Royden} it is therefore also \emph{compact}. The definitions of each of these terms can be found in any standard topology reference and we point the reader to \cite{Royden} for further details.

  \begin{definition}[Upper Semi-Continuous] 
    A compact-valued correspondence $g: A \rightarrow B$ is
    \emph{upper semi-continuous} \cite{Findforthis} at a point $a \in A$ if $g(a)$ is non-empty
    and if, for every sequence $a_n \rightarrow a$ and every sequence
    $(b_n)$ such that $b_n \in g(a_n)$ for all $n$, there exists a
    convergent subsequence of $(b_n)$ whose limit point $b$ is in
    $g(a)$.
  \end{definition}

  \fb{perhaps we have to introduce some terminology from topology to keep the results self-contained?}

  \begin{theorem}[Kakutani] 
    Let $A$ be a compact and convex in a Euclidean space and $g: A \rightarrow \mathbf{P}(A)$ (where
    $\mathbf{P}(A)$ denotes the power set of A) be upper semi-continuous, with nonempty, convex and compact values.
    Then $g$ has a fixed point, that is a point $\bar{a}$ such that $\bar{a} \in g(\bar{a})$.
    \cite{Kakutani}
  \end{theorem}

  Note that, when acting on a simplex, the function $g: \Delta \rightarrow \textbf{P}(\Delta)$, where $\textbf{P}(\Delta)$ denotes all nonempty, closed and convex subsets of $\Delta$ only has to satisfy the upper-semi continuity condition to admit a fixed point.

	\ah{Probably need a quick intro into why the argmax is closed, convex and nonempty.}

@@

\subsection{Existence of NA-CTFP}

@@

\subsection{Convergence of NA-CTFP}

@@

  with no alteration to $A$. The modification itself is small, however it results in the zero-sum
  assumption being violated. With the same choices of $\beta$ and $w$, this results in the periodic
  orbit seen in Figures \ref{fig::nonconvergentShapley} and \ref{fig::3PlayerChainNoNoise}. Here,
  the orbits reach a stable limit cycle which to be centred around the interior NE.

@@

  \begin{figure}[t]
    \centering
    \includegraphics[width = 0.7 \textwidth]{Figures/3PlayerChainNoNoise.png}
    \caption{\label{fig::3PlayerChainNoNoise} Distance to the Nash Equilibrium strategy along orbits
    of FP showing oscillations which are bounded away from the NE (i.e. they do not cross the NE).
    The x-axis shows logarithmic time since FP includes a factor of $1/t$, which has the effect of
    slowing down the orbit over time.}
  \end{figure}

@@

  \begin{figure}[t]
    \centering
    \includegraphics[width = \columnwidth]{Figures/Noise10Player.png}
    \caption{\label{fig::Noise10Player} Trajectories of NA-CTFP in a 10 player game with additive noise (Left) No noise is introduced and learning converges directly to an NE. (Middle) $\gamma = 0.1$, the trajectories converge to a fixed point but removed from the NE. (Right) $\gamma = 0.5$, the trajectories converge to a fixed point which is even further away from the NE.}
  \end{figure}

@@

  In Figure \ref{fig::Noise10Player}, we consider a zero-sum NA game with 10 players. When there is
  no noise, it can be seen that FP reaches a fixed point which, since we set $w^{\mu \mu} = 0$,
  corresponds to an NE. After increasing $\gamma$, however, we find that the agents no longer
  converge to this NE, but rather shift away from it. What is interesting, however, is that the
  orbits do still reach a stationary state in the long run which suggests that FP is still able to
  converge with the introduction of noise. Figure \ref{fig::Noise20Player} shows similar behaviour
  for $N = 20$, and in our experiments we found this to be ubiquitous regardless of the choice of
  $N$ or $n$.

  In Figure \ref{fig::3PlayerChainNoise} we revisit the Three Player Chain of Section
  \ref{sec::NonConv}, now under the influence of additive noise. For the sake of brevity, we only
  display the distance to the Nash Equilibrium of the first player's action, since the other agents
  behave in the same way. It can be seen that a small amount of noise has the effect of decreasing
  the size of the periodic orbit (c.f. Figure \ref{fig::3PlayerChainNoNoise}). However, as $\gamma$
  is increased to $0.5$, the algorithm seems to exhibit convergence to the NE. The implication is
  that the addition of noise may cause periodic behaviour to break and lead to the Nash Equilibrium.
  An interesting point to note is that this behaviour is in stark contrast to the replicator
  dynamic (RD) \cite{Maynard-Smith}, another adaptive algorithm linked to multi agent learning
  \cite{CyclesAdversarialLearning}. In \cite{Imhof} and \cite{Galla}, it was found that the
  introduction of random mutations can remove convergent behaviour and instead lead to periodicity.

>>>>

\title{Equilibria and Convergence of Fictitious Play on Network Aggregative Games}

@@

\begin{abstract}
  Understanding the long term behaviour of learning algorithms is an important object of study if we are to design systems which can be considered safe. However, as the number of players in a game increase, it is known that the dynamics of learning becomes increasingly chaotic. As such, the most rigorous results of learning dynamics are limited to two player games.

  In this study, we contribute towards the study of learning with multiple agents. We do this through the introduction of learning on network aggregative games, in which each player's reward is dependent on its own strategy and a convex combination of its neighbours. In particular, we present a continuous time analysis of the Fictitious Play learning dynamic on such a game. We reach a number of conclusions. In particular, we show that Fictitious Play reaches a fixed point when the game is zero-sum and provide conditions under which this fixed point corresponds to a Nash Equilibrium. In addition, we show that agents learning through Fictitious Play achieve no-regret, regardless of the choice of game. Finally, we present experimental evidence of a family of games for which Fictitious Play reaches a limit cycle and evidence that the introduction of noise has the potential to break this cyclic behaviour and allow agents to reach the Nash Equilibrium.
\end{abstract}

@@

Multi Agent Learning \cite{Schwartz} requires a number of agents to adapt in an environment where
each agent must adapt in response to the behaviour of the other agents. This leads to a
fundamentally non-stationary problem, which presents a challenge to designing effective learning
policies. Even when there are a small number of agents in the game, learning has been shown to lead
to non-stationary, and even chaotic behaviour \cite{SatoChaos}, and this problem becomes even more
pronounced as the number of agents increase \cite{Sanders}. In this light, it would seem that, when there is a large population of agents, it is almost impossible to understand the long term behaviour of multi agent learning \cite{PiliourasChaoticMaps}. This, of course, poses a problem for AI safety, for which it is necessary that learning systems are designed so that their ultimate behaviour is guaranteed to satisfy some predetermined goals.

To deal with this problem, one must attempt to reduce the many player game to something which is
tractable. A number of reductions have been proposed, most notably \emph{Mean Field} Games
and \emph{Aggregative} Games. The former makes the assumption of an infinite
number of agents, so that the population can be represented through a distribution over players'
states \cite{CainesPaper} . Every agent then updates their action profiles depending on this
distribution. In the latter each agent plays considers a real valued function which is a convex
combination of the states of the other agents. In both formulations, a many player game is reduced
to the set of two player games which each agent is involved in.

Both formulations have been the object of rigorous study and it has been shown that agents reach an equilibrium when learning on such games (c.f. \cite{MFGLearningPapers} for Mean Field Games and
\cite{Aggregative Papers} for Aggregative Games). However, both also present a fundamental limitation.
Namely, they both require that agents have access to the action profiles of the entire population.
This could be through communication with all other agents, or through the intervention of a central
coordinator who is able to the entire population. Whilst recent work aims to relax this assumption
through the introduction of noise \cite{MFG-FP} or partial observability \cite{AAMASPaper}, the
requirement that each agent updates their actions based on the entire population is rather strong.

In this study, we investigate a variant of the aggregative game known as the \emph{Network
Aggregative} (NA) Game. This format assumes that there is an underlying communication network
through which agents interact. Then each agent updates their actions according only to those agents
with whom they communicate. This significantly relaxes the communication load on each agent and
lifts the need for a central coordinator. Recent work on the NA game has shown that it is possible
for agents to reach an equilibrium strategy in an entirely distributed manner \cite{Grammatico,
LeaderFollower, MyopicAgents}. We contribute in this direction by analysing the long term behaviour of learning on an NA game. In particular, we analyse the Fictitious Play Learning
Algorithm \cite{Brown, Harris}, in which agents are assumed to be myopic, in that they react solely
to the previous behaviour of the others.

@@

  The main contribution of this work is to introduce learning on the Network Aggregative Game
  through the action of Fictitious Play (FP).

  We first show that a Nash Equilibrium exists and that FP admits
  solutions in this setting. In particular, we study zero-sum games and show that
  FP converges to a fixed point, which for a network without any
  self-loops corresponds to a Nash Equilibrium. In addition, we find that, for games
  which are not zero-sum, agents following FP are able to
  achieve no regret.

  Finally, we explore FP through numerical simulations to consider the
  question of whether it always converges. We answer this in the negative by finding a family of
  games in which action profiles cycle around the Nash Equilibrium. Finally, our experiments
  document how noise affects the convergence of FP. Our experiments suggest that, under
  the presence of noise, the algorithm still reaches a fixed point, but perhaps not the Nash
  Equilibrium. This presents an interesting avenue for future research.

@@

The NA game is a recent introduction \cite{Parise2015} which presents a variant of aggregative games
by adding an underlying structure to the population. Since its introduction, distributed algorithms
have been built with the aim of finding NE. In particular, \cite{Parise2015, Parise2020} consider
the case in which payoffs are given by Lipschitz functions with unique minimisers and apply standard
topological fixed point towards designing algorithms which will converge to the NE. We note that the
\emph{Mann Iteration}, one of the algorithms applied in \cite{Parise2020} shares a remarkably
similar form with the discrete variant of FP. To the best of our knowledge, this link has not been
explored in the literature and is an avenue for research which we are currently pursuing. Another
common algorithm for distributed NE seeking is the projected gradient (resp. subgradient) dynamics
which is explored in \cite{Zhang2020} (resp. \cite{Shokri2020, Shokri2021}). In all of these works,
the cost function is assumed to be convex, and therefore have a unique minimiser. In fact this is a
common assumption in works concerning NA Games \cite{Zhu2021, Lei2020}, which we believe is due to
its ubiquity in control settings. We have not yet come across works which consider NA games from the
point of view of payoff matrices, which are more common in multi agent learning settings.
Furthermore, to the best of our knowledge, this is the first work which introduces the application
of a strict learning algorithm in the NA setting.

@@

  In this section we present our main results. First, we establish the existence of a Nash Equilibrium in the NA game, as in Def. \refname{def::NE}. Then we show that the Fictitious Play dynamic on the NA game admits solutions (i.e. the NA-CTFP exists). Finally, we show that NA-CTFP reaches a fixed point when the NA game is zero-sum and that, when the network has no self-loops (i.e. is simple), NA-CTFP reaches a Nash Equilibrium. For the sake of brevity, we defer the proofs of our statements, as well as the standard topological arguments used to derive them to the Supplementary Material (Sections (S4 - S6)).

@@

  With these in place, we can state our main results.

@@

@@

@@

  with no alteration to $A$. The modification itself is small, however it results in the zero-sum
  assumption being violated. With the same choices of $\beta$ and $w$, this results in the periodic
  orbit seen in Figures \ref{fig::nonconvergentShapley}. Here,
  the orbits reach a stable limit cycle which to be centred around the interior NE.

@@

@@

@@

  In Figure \ref{fig::Noise20Player}, we consider a zero-sum NA game with 20 players. When there is
  no noise, it can be seen that FP reaches a fixed point which, since we set $w^{\mu \mu} = 0$,
  corresponds to an NE. After increasing $\gamma$, however, we find that the agents no longer
  converge to this NE, but rather shift away from it. What is interesting, however, is that the
  orbits do still reach a stationary state in the long run which suggests that FP is still able to
  converge with the introduction of noise.

  In Figure \ref{fig::3PlayerChainNoise} we revisit the Three Player Chain of Section
  \ref{sec::NonConv}, now under the influence of additive noise. For the sake of brevity, we only
  display the distance to the Nash Equilibrium of the first player's action, since the other agents
  behave in the same way. It can be seen that a small amount of noise has the effect of decreasing
  the size of the periodic orbit. However, as $\gamma$
  is increased to $0.5$, the algorithm seems to exhibit convergence to the NE. The implication is
  that the addition of noise may cause periodic behaviour to break and lead to the Nash Equilibrium.
  An interesting point to note is that this behaviour is in stark contrast to the replicator
  dynamic (RD) \cite{Maynard-Smith}, another adaptive algorithm linked to multi agent learning
  \cite{CyclesAdversarialLearning}. In \cite{Imhof} and \cite{Galla}, it was found that the
  introduction of random mutations can remove convergent behaviour and instead lead to periodicity.
